Examining equity in accessibility to bike share - a balanced floating catchment area approach

### 2021-06-16 Bike-share-spatial-equity/Bike-share-spatial-equity.Rmd
Revisions with Elise

```{r load-packages, include=FALSE}
library(biscale)
library(cowplot)
library(data.table)
library(disk.frame)
library(dplyr)
library(gdistance)
library(gridExtra)
library(kableExtra)
library(pycno)
library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.htmllibrary(readr)
library(raster)
library(rgdal)
library(sf)
library(sobiEquity) # Data package
library(tidyverse)
library(tinytex)
library(units)
```

@@

As of June 2020, HBS has 900 bikes, 130 stations [see Figure \ref{fig:sobi-stations-in-hamilton}], and over 26,000 active memberships [@hamiltonHamiltonBikeShare2015]. The core service area remains 40 sq.km. The program has twelve "equity" stations and 118 "conventional" stations (i.e., that were not explicitly added to address inequities in the program). The City of Hamilton has positioned stations between 300 and 600 m apart [@scottRouteChoiceBike2021], but anticipates that they will service those living within a 250 m buffer [@hamiltonsobi2015]. The latter constitutes a normative statement: people ought to be able to access a station in less than 600 m if they live in the core service area, with most usage coming from 250 m around. However, it is not known how far people are actually willing to travel to reach a station. It would be reasonable to assume that people are willing to walk beyond this threshold to access other stations if the ones nearest them have no supply of bicycles.

```{r hamilton-and-sobi-service-area, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "The core service area of Hamilton Bike Share is outlined in black. Hamilton Census Metropolitan Area is shown in grey."}
ggplot() + 
  geom_sf(data = hamilton_da_2016,
          color = NA,
          fill = "gray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "black") + 
  theme(panel.background = element_rect(fill = "white"))
```

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "100%", fig.cap = "The spatial distribution of bicycle share stations in Hamilton, Ontario. Conventional stations are in blue and equity stations are in green. The service area of Hamilton Bike Share is outlined in black and the city's downtown core is outlined in pink. Hamilton Census Metropolitan Area is shown in grey."}
ggplot() + 
  geom_sf(data = cropped_da,
          color = NA,
          fill = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          size = 1,
          color = "pink") +
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_type == "Conventional" & 
                            hub_status == "Active"), # Remove deactivated hubs
          aes(size = RACKS_AMOU),
          shape = 21,
          fill = "blue") +
  geom_sf(data =  sobi_hubs %>% 
            dplyr::filter(hub_type == "ERI"), 
          aes(size = RACKS_AMOU), 
          shape = 21,
          fill = "green") +
  scale_size(name = "Number of racks", range = c(0.5, 4)) +
  theme(panel.background = element_rect(fill = "white"))
```

@@

The first step in the FCA method is to allocate the population to be serviced by each docking station:
\begin{equation}
\label{eq:population-allocation}
P_j = {\sum_{i = 1}^{n} P_i{w_{ij}}}
\end{equation}

@@

```{r da-population, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "Population in all dissemination areas (outlined in white) that are inside or touch the bounding box of the Hamilton Bike Share core service area (outlined in black)."}
ggplot() +
  geom_sf(data = hamilton_da_2016 %>%
            dplyr::filter(sobi_service == "Yes"),
          aes(fill = population),
          color = "white") +
  geom_sf(data = sobi_service,
          fill = NA) +
  scale_fill_distiller(palette = "YlOrRd", 
                       direction = 1) +
  theme(panel.background = element_rect(fill = "white"))
```

```{r interpolated-population, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "Interpolated population in the Hamilton Bike Share's core service area (outlined in black) and within 30 minutes of walking to the core service area. Each population cell is 50-by-50 m in size. The downtown area is outlined in dark green."}
ggplot() + 
  geom_sf(data = cropped_da,
          fill = "gray",
          color = NA) + 
  #geom_sf(data = population_50x50,
  #        aes(color = population)) +
   #scale_fill_gradient(colours = "YlOrRd") +
  geom_tile(data = cbind(st_drop_geometry(population_50x50), 
                         st_coordinates(population_50x50)),
            aes(x = X, 
                y = Y, 
                fill = population)) +
  scale_fill_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  theme(panel.background = element_rect(fill = "white"))
```

@@

With a walking distance of three minutes, we found that there are 25.2 bicycle racks per person system-wide in the original system configuration (i.e., without equity stations). The addition of equity stations increases this ratio slightly to 25.4 bicycles per person. Figure \ref{fig:figure-6} presents a comparison of accessibility between the systems for the population cells. Accessibility is fairly uniform overall, with the exception of two small areas where accessibility is slightly higher. This high level of system-wide accessibility occurs because the population that can reach bicycle share stations when travel time is 3 minutes or less is very limited, and accessibility is strongly shaped by a few locations that concentrate population and stations.

```{r figure-6, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "Accessibility at 3 minutes walk (minimum threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 3 minutes walk
accessibility_3min <- accessibility %>% 
  dplyr::filter(threshold == "3 min")

ggplot() +
  geom_sf(data = population_50x50,
          color = "gray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "3 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_3min), st_coordinates(accessibility_3min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

@@

ggplot() +
  geom_sf(data = population_50x50,
          color = "gray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "5 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_5min), st_coordinates(accessibility_5min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

@@

ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "10 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_10min), st_coordinates(accessibility_10min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

@@

ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "15 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_15min), st_coordinates(accessibility_15min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA) +

  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

@@

To highlight the benefit of using the BFCA approach, we conducted a comparative analysis with the 2SFCA method. The latter method does not normalize the impedance to allocate population and level of service proportionally.

@@

```{r conventional-5-min-sum-non-equity}
sum(conventional_5min_non_equity$los$los)
sum(conventional_5min_non_equity$accessibility$accessibility)
```

```{r conventional-5-min-sum-all}
sum(conventional_5min_non_equity$los$los)
sum(conventional_5min_non_equity$accessibility$accessibility)
```

@@

A unique property of the balanced FCA method is that data can be reaggregated while preserving the total population in the core service area and the supply at each docking station. This avoids demand and supply inflation, and also enables us to present findings in a way that is perhaps more intuitive to interpret. Therefore, we reaggregated population and accessibility from the 50-by-50 m population cells to the dissemination area. We then can compared accessibility to median total household income statistics from the 2016 Canadian Census.

@@

To further explore this issue and examine vertical equity, we aggregated accessibility by median total household income across all geographies (see Table \ref{tab:accessibility-income}). Similar to Hosford and Winters  [-@hosfordWhoArePublic2018], we found that Hamilton's PBSP already serviced a large proportion of the population in the bottom and second quintiles of median total household income. This may well be an artifact of the spatial socioeconomic and demographic profile of Hamilton, where the most dense parts of the city (where a PBSP is most easily launched) are also those with relatively lower incomes. On the other hand, the levels of accessibility to Hamilton's PBSP are generally lower for populations in the bottom 20% of median total household income, compared to populations in the top 20%.

@@

It does not have the property of preserving these quantities. This produces a There is an inflation/deflection effect

>>>>

```{r load-packages, include=FALSE}
library(biscale)
library(cowplot)
library(data.table)
library(disk.frame)
library(dplyr)
library(gdistance)
library(ggmap)
library(ggspatial)
library(gridExtra)
library(kableExtra)
library(pycno)
library(r5r) # the r5r package requires Java Development Kit version 11, which can be downloaded from https://www.oracle.com/java/technologies/javase-jdk11-downloads.htmllibrary(readr)
library(raster)
library(rgdal)
library(sf)
library(sobiEquity) # Data package
library(tidyverse)
library(tinytex)
library(units)
```

@@

As of June 2020, HBS has 900 bikes, 130 stations [see Figure \ref{fig:sobi-stations-in-hamilton}], and over 26,000 active memberships [@hamiltonHamiltonBikeShare2015]. The core service area remains 40 sq.km. The program has twelve "equity" stations and 118 "conventional" stations (i.e., that were not explicitly added to address inequities in the program). The City of Hamilton has positioned stations between 300 and 600 m apart [@scottRouteChoiceBike2021], but anticipates that they will service those living within a 250 m buffer [@hamiltonsobi2015]. The latter constitutes a normative statement: people ought to be able to access a station in less than 600 m if they live in the core service area, with most usage coming from 250 m around. However, it is not known how far people are actually willing to travel to reach a station. It would be reasonable to assume that people are willing to walk beyond this threshold to access other stations if the ones nearest them have no supply of bicycles.
```{r}

```

```{r hamilton-and-sobi-service-area, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "The core service area of Hamilton Bike Share is outlined in blue. Hamilton Census Metropolitan Area is shown in grey."}
geo_context <- ggplot() + 
  geom_sf(data = hamilton_da_2016,
          color = NA,
          fill = "gray") + 
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "blue") + 
  theme(panel.background = element_rect(fill = "white"),
        axis.text = element_blank(),
        axis.ticks = element_blank()) +
  #annotate("text", x = -80, y = 43.5, label = "Lake Ontario") +
  annotation_scale(location = "bl", width_hint = 0.4) +
  annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.25, "in"), pad_y = unit(0.25, "in"),
        style = north_arrow_fancy_orienteering)

# Inset map
geo_context_inset <- get_stamenmap(bbox = c(left = -79,
                                    bottom = 43,
                                    right = -81,
                                    top = 44),
          maptype = "toner-lite", 
          crop = FALSE,
          zoom = 8)
# plot map
geo_context_inset <- ggmap(geo_context_inset)  +
    coord_sf(crs = st_crs(4326))  +
    #xlab("Longitude") + ylab("Latitude") +
  annotation_scale(location = "bl", width_hint = 0.4) +
  theme_minimal() +
  theme(axis.text = element_blank(),
        panel.border = element_rect(color = "black",
                                    fill = NA),
        axis.title = element_blank())

#geo_context + geo_context_inset
ggdraw() + 
  draw_plot(geo_context, 0.0, 0.0, 0.8, 1) + 
  draw_plot(geo_context_inset, 0.5, 0.4, 0.6, 0.6) +
  draw_plot_label(x = c(.25,.5), y = c(0.45,.6), 
                  label =  c("Hamilton Bikeshare \nCore Service Area",  "Lake Ontario"),
                  hjust = 0, 
                  size = 8) +
  draw_line(x = c(0.34, 0.4),
    y = c(0.43, 0.5),
    color = "blue", 
    size = 1,
    lineend = "round",
    linejoin = "mitre",
    arrow = arrow(length = unit(0.1, "inches")))
```

```{r sobi-stations-in-hamilton, echo=FALSE, fig.align = 'center', out.width = "100%", fig.cap = "The spatial distribution of bicycle share stations in Hamilton, Ontario. Conventional stations are in blue and equity stations are in green. The service area of Hamilton Bike Share is outlined in black and the city's downtown core is outlined in pink. Hamilton Census Metropolitan Area is shown in grey."}
ggplot() + 
  geom_sf(data = cropped_da,
          color = NA,
          fill = "gray") +
  geom_sf(data = sobi_service, 
          fill = NA,
          color = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          size = 1,
          color = "pink") +
  geom_sf(data = sobi_hubs %>% 
            dplyr::filter(hub_status == "Active"), # Remove deactivated hubs
          aes(size = RACKS_AMOU, fill = hub_type),
          shape = 21) +
  scale_size(name = "Number of racks", 
             range = c(0.5, 3.5)) +
  scale_fill_manual(name = "Hub Type",
                    values = c("Conventional" = "blue", "ERI" = "green")) +
  theme(panel.background = element_rect(fill = "white"),
        axis.text = element_blank(),
        legend.position = "bottom",
        axis.ticks = element_blank()) +
  annotation_scale(location = "bl", width_hint = 0.4,
                   pad_x = unit(0.4, "in"), pad_y = unit(0.2, "in"),) +
  annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.25, "in"), pad_y = unit(0.20, "in"),
        style = north_arrow_fancy_orienteering)
```

@@

The first step in the FCA method is to allocate the population to be serviced by each docking station:

\begin{equation}
\label{eq:population-allocation}
P_j = {\sum_{i = 1}^{n} P_i{w_{ij}}}
\end{equation}

@@

```{r interpolated-population, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "Interpolated population in the Hamilton Bike Share's core service area (outlined in black) and within 30 minutes of walking to the core service area. Each population cell is 50-by-50 m in size. The downtown area is outlined in dark green."}
ggplot() + 
  geom_sf(data = cropped_da,
          fill = "gray",
          color = NA) + 
  #geom_sf(data = population_50x50,
  #        aes(color = population)) +
   #scale_fill_gradient(colours = "YlOrRd") +
  geom_tile(data = cbind(st_drop_geometry(population_50x50), 
                         st_coordinates(population_50x50)),
            aes(x = X, 
                y = Y, 
                fill = population)) +
  scale_fill_distiller(palette = "YlOrRd", 
                        direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA,
          colour = "black") +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()) +
  annotation_scale(location = "bl", width_hint = 0.4,
                   pad_x = unit(0.4, "in"), pad_y = unit(0.2, "in"),) +
  annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.25, "in"), pad_y = unit(0.20, "in"),
        style = north_arrow_fancy_orienteering)
```

@@

With a walking distance of three minutes, we found that there are 25.2 bicycle racks per person system-wide in the original system configuration (i.e., without equity stations). The addition of equity stations increases this ratio slightly to 25.4 bicycles per person. Accessibility is fairly uniform overall, with the exception of two small areas where accessibility is slightly higher. This high level of system-wide accessibility occurs because the population that can reach bicycle share stations when travel time is 3 minutes or less is very limited, and accessibility is strongly shaped by a few locations that concentrate population and stations. For this reason, accessibility is highly concentrated in small areas around the stations. The map is not shown but can be recreated using the source code and data.

```{r figure-6, eval = FALSE, echo=FALSE, fig.align = 'center', out.width = "90%", fig.cap = "Accessibility at 3 minutes walk (minimum threshold) compared between current system with equity stations and the original system without equity stations."}
#accessibility plot for current system with equity stations at 3 minutes walk
accessibility_3min <- accessibility %>% 
  dplyr::filter(threshold == "3 min")

ggplot() +
  geom_sf(data = population_50x50,
          color = "gray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "3 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_3min), st_coordinates(accessibility_3min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", direction = 1) +
  geom_sf(data = sobi_service,
          fill = NA) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())  +
  annotation_scale(location = "bl", width_hint = 0.4,
                   pad_x = unit(0.4, "in"), pad_y = unit(0.2, "in"),) +
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
        style = north_arrow_fancy_orienteering)

```

@@

ggplot() +
  geom_sf(data = population_50x50,
          color = "gray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "5 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_5min), st_coordinates(accessibility_5min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", 
                       direction = 1,
                       trans = "log",
                       breaks = c(0.001, 0.02, 0.4, 8)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())  +
  annotation_scale(location = "bl", width_hint = 0.4,
                   pad_x = unit(0.4, "in"), pad_y = unit(0.2, "in"),) +
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
        style = north_arrow_fancy_orienteering)
```

@@

ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "10 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_10min), st_coordinates(accessibility_10min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd",
                       direction = 1,
                       trans = "log",
                       breaks = c(0.00004, 0.0003, 0.002)) +
  geom_sf(data = sobi_service,
          fill = NA) +
  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())  +
  annotation_scale(location = "bl", width_hint = 0.4,
                   pad_x = unit(0.4, "in"), pad_y = unit(0.2, "in"),) +
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
        style = north_arrow_fancy_orienteering)
```

@@

ggplot() +
  geom_sf(data = population_50x50,
          color = "lightgray") +
  #geom_sf(data = accessibility %>%
  #          dplyr::filter(threshold == "15 min"),
  #        aes(color = accessibility)) +
  #scale_color_distiller(palette = "PuRd", 
  #                      direction = 1) +
  geom_tile(data = cbind(st_drop_geometry(accessibility_15min), st_coordinates(accessibility_15min)),
            aes(x = X, y = Y, fill = accessibility)) +
  scale_fill_distiller(palette = "PuRd", 
                       direction = 1,
                       trans = "log",
                       breaks = c(0.000006, 0.00005, 0.0003)) +
  geom_sf(data = sobi_service,
          fill = NA) +

  facet_grid(system ~ threshold)  +
  geom_sf(data = community_downtown %>%
            dplyr::filter(NAME == "Hamilton"),
          fill = NA,
          color = "dark green") +
  # geom_sf(data =  sobi_hubs %>% 
  #           dplyr::filter(hub_type == "ERI"), 
  #         size = 1.5, shape = 21, 
  #         fill = "green") +
  theme(panel.background = element_rect(fill = "white"),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())  +
  annotation_scale(location = "bl", width_hint = 0.4,
                   pad_x = unit(0.4, "in"), pad_y = unit(0.2, "in"),) +
  annotation_north_arrow(location = "tr", which_north = "true", 
        pad_x = unit(0.0, "in"), pad_y = unit(0.0, "in"),
        style = north_arrow_fancy_orienteering)
```

@@

To highlight the benefit of using the BFCA approach, we conducted a comparative analysis with the 2SFCA method. The latter method does not allocate population and level of service proportionally, which means that there are often substantial amounts of multiple counting each. Figure X shows for the 10-minute threshold the level of inflation/deflation that results when the balanced and the conventional approaches are compared. The inflation/deflation is defined as the ratio of the level of service and accessibility calculated using the conventional approach to the balanced approach. This effect has been documented in PAEZ-HIGGINS-VIVONA and Chen?. Inspection of Fig. REFERENCE TO FIGURE, shows that in some locations, the level of service is deflated by up to 40\% (or whatever) and accessibility is inflated by as much as 40\% (or whatever).

@@

@@

Next, we wish to examine accessibility differences by income levels. To this end we take our micro-zones and impute the median household income of their corresponding DA. Use of bivariate maps helps us to visualize the level of accessibility by level of income. Median total household income statistics were drawn from the 2016 Canadian Census.

@@

To examine vertical equity we need to estimate what level of accessibility accrues to how many people in different income strata. A unique property of the balanced FCA method (which does not hold for the conventional approach because of the inflation/deflation issues discussed above) is that data can be re-aggregated accessibility and population using income as an aggregator criterion, while preserving the total population and the level of service. This avoids demand and supply inflation, and also enables us to present findings in a way that is perhaps more intuitive to interpret. Therefore, we aggregate the accessibility and population by the median household income as imputed from the DAs (see Table \ref{tab:accessibility-income}). Similar to Hosford and Winters  [-@hosfordWhoArePublic2018], we found that Hamilton's PBSP already serviced a large proportion of the population in the bottom and second quintiles of median total household income. This may well be an artifact of the spatial socioeconomic and demographic profile of Hamilton, where the most dense parts of the city (where a PBSP is most easily launched) are also those with relatively lower incomes. On the other hand, the levels of accessibility to Hamilton's PBSP are generally lower for populations in the bottom 20% of median total household income, compared to populations in the top 20%.

@@
